\subsection{Compound Types}
\label{sec:compound-types}

Compound types are constructions that can be made up of the primitive
types. In \Trilogy{}, all compound types are heterogeneous - they may
contain many values of different types.

Similarly to primitives, compound types have string and source
representations. The string representation is computed by recursively
converting component values into their string representations. The
source representation is once again the same as you would write it
in source code.

\subsubsection{Struct}

A Struct wraps a single other value with a label. The label looks a
lot like an Atom, and the value can be any single value (including
a compound value). The Struct type is a structural and non-reference
type.

Each distinct label for a Struct actually acts more like its own ``type'' from
a type-system perspective, since labels cannot be referenced dynamically, only
in literal patterns. In a pattern matching context, you could think of them as
sum-type variants, but without defining them as a members of a specific list
of variants.

Given that, in documentation we rarely refer to the Struct type as ``Struct'',
but instead using the exact label that is expected, as in \texttt{'label(T)}

The string representation of a Struct is the label (without the single quote)
followed by the string representation of the value in parentheses.

\begin{table}[H]
    \centering
    \begin{tabular}{ll}
        \hline
        \textbf{Source} & \textbf{String} \\
        \hline
        \texttt{'hello("world")} & \texttt{hello(world)} \\
        \texttt{'number(3)} & \texttt{number(3)} \\
        \texttt{'compound([1, 2, 3])} & \texttt{compound([1, 2, 3])} \\
        \hline
    \end{tabular}
\end{table}

\subsubsection{Tuple}

A Tuple attaches a pair of values to make a single compound value, and is written
with the colon or ``cons'' operator (\op{:}). The Tuple type is a structural and
reference type. In type annotation, the Tuple type is written as in \texttt{T : U}.

While a Tuple can only contain two values, one of those values may be another
Tuple to emulate larger collections. When chained together, the cons operator
associates right.

While finite Tuples are useful for representing structured data, recursive Tuples
are useful as a representation of (linked) lists. In this interpretation, we use
\kw{unit} to indicate the empty List, and any value consed with a List on the
right side to be a List. Lists may be consumed recursively. We use the name
List (and type annotation \ty{List T}) to refer to constructions of Tuples in
this way, but List is not truly a distinct ``type''; it's simply a special case
of the Tuple type.

The string representation of a Tuple consists of the string representations of its
two parts, separated by a colon much like its source representation.

\begin{table}[H]
    \centering
    \begin{tabular}{ll}
        \hline
        \textbf{Source} & \textbf{String} \\
        \hline
        \texttt{"hello" : "world"} & \texttt{hello : world} \\
        \texttt{3 : true} & \texttt{3 : true} \\
        \texttt{1 : 2 : 3 : unit} & \texttt{1 : 2 : 3 : unit} \\
        \hline
    \end{tabular}
\end{table}

\subsubsection{Array}

An Array is a contiguous, ordered sequence of values, backed by an array of
the implementation language. In terms of function and capabilities, Arrays
and Lists are pretty much equivalent, but the performance characteristics
of the two are quite distinct, particularly when the list will be resizing
frequently, versus having frequent random access. The Array type is a structural
and reference type. In type annotation, the Array type is written \ty{Array T}.

Array literals are written with square brackets, and items separated by
comma. In Array construction, the spread operator (\op{..}) can be used
to flatten one Array into the new one. In Array decconstruction, the spread
operator may be used \emph{at most once} to refer to the rest of the elements
not already matched explicitly by the pattern.

Typical array operations include direct access using the indexing operator,
push (\texttt{+}), and concatenation or glue (\texttt{<>}).

The string representation of an Array is simply the string representation of
each of its elements, separated by commas within square brackets.

\begin{table}[H]
    \centering
    \begin{tabular}{ll}
        \hline
        \textbf{Source} & \textbf{String} \\
        \hline
        \texttt{[1, 2, 3]} & \texttt{[1, 2, 3]} \\
        \texttt{["hello", 3, "world"]} & \texttt{[hello, 3, world]} \\
        \texttt{[1, 2, ..[3, 4], 5, 6]} & \texttt{[1, 2, 3, 4, 5, 6]} \\
        \hline
    \end{tabular}
\end{table}

\subsubsection{Set}

A Set is an unordered collection of values containing no duplicates. Though
values of any type may be contained in a Set, duplicates are compared as if
by the structural equality operator (\op{eq}). The Set type is a structural
and reference type. In type annotation, the Set type is written \ty{Set T}

Sets come with useful operations union (\op{|}), intersection (\op{\&}),
add (\op{+}), remove (\op{-}), difference (\op{/}), and symmetric
difference (\op{\textasciicircum}).

Set literals are written with curly braces and items separated by comma.
In pattern matching, the spread operator may be used to refer to the elements
not already matched explicitly by the pattern. In Set construction, the
spread operator is not supported; prefer to use the Set operators instead.

The string representation of a Set is the string representation of
each of its elements, separated by commas within curly braces.

\begin{table}[H]
    \centering
    \begin{tabular}{ll}
        \hline
        \textbf{Source} & \textbf{String} \\
        \hline
        \texttt{\{1, 2, 3\}} & \texttt{\{1, 2, 3\}} \\
        \texttt{\{"hello", "hello", "world"\}} & \texttt{\{hello, world\}} \\
        \hline
    \end{tabular}
\end{table}

\subsubsection{Record}

A Record is an unordered collection of key-value pairs, often also called
a dictionary, map, or object. Record keys are unique. The Record type is a
structural and reference type.

In type annotation, the record type is written \ty{Record K V} where keys of
type \ty{K} refer to values of type \ty{V}. As a special case, combining
multiple Record types using the type product operator (\op{\&}) allows a Record
to specify specific value types for different key types, e.g.
\ty{Record K V \& Record L W}.

The main operation of a Record is the indexing operator. The keys of a
Record may be any value; like with Sets, the keys are compared as if by
the structural equality operator (\op{==})

A Record literal is written as comma separated \texttt{key: value} pairs
within curly braces. In Record construction, the spread operator (\op{..})
may be used to combine two or more Records. In pattern matching, the spread
operator may be used once to refer to the rest of the entries not already
matched explicitly by the pattern.

Though Records are unordered, keys specified ``later'' in the source file
(whether literally or via a spread) take precedence over keys specified
earlier.

The string representation of a Record is its stringified \texttt{key: value}
pairs in curly braces.

\begin{table}[H]
    \centering
    \begin{tabular}{ll}
        \hline
        \textbf{Source} & \textbf{String} \\
        \hline
        \texttt{\{1: "one", 2: "two"\}} & \texttt{\{1: one, 2: two\}} \\
        \texttt{\{"hello": 1, "hello": 2, "world": 3\}} & \texttt{\{hello: 2, world: 3\}} \\
        \hline
    \end{tabular}
\end{table}

\subsubsection{Iterator}
\label{sec:iterators}

Iterators are the result of certain functions; it cannot be specified literally.
Iterators act as ordered sequences that can be ``iterated'' over, yielding its
elements one by one into a pattern. Iteration is performed lazily, therefore
infinite iterators are possible. The Iterator type is a non-structural and
reference type. In type annotation, the Iterator type is written \ty{Iter T}.

Iterators in \Trilogy{} are \emph{immutable}, and are typically interacted
with via an iterative context (e.g. a \kw{for} loop or comprehension), but
may also be manually stepped through by unifying as if it were a List. Their
being immutable means that if a previous state of an iterator is stored and then
re-iterated over, that iterator will produce the same values. Internally, they
are simply cached.

Important to note, however, is that an iterators internal cache is \emph{not}
shared between parallel executions. Of particular relevance for iterators which
are non-deterministic or perform side effects (such as an iterator that reads
its values via I/O) will potentially have different results in different executions,
and may cause more side effects than expected to occur. Using iterators to
represent side effects can be convenient, but to use such iterators in multiple
executions is not recommended as it is most likely just confusing.

Also important to note is that an iterator's cache will grow in size unless
references to its past states are dropped. For finite or deterministic iterators
this is not such a big deal (likely the elements are already stored in a backing
collection) but for infinite iterators this can lead to an infinitely large
storage. Do be careful when retaining references to particular states of iterators.

As there is currently no way to specify a literal Iterator directly, there is no
source representation of the Iterator type. Additionally, since an Iterator
may be infinite and may cause side effects, its string representation is simply
defined to be the literal string \texttt{<iterator>}.
