\section{Type System}
\label{sec:types}

\Trilogy{} is (currently) a dynamically typed language; all types are determined
and checked at runtime. Recall that ``dynamically typed'' is not the absence of
a type system, but the presence of a type system so complex that it cannot be
fully statically analyzed.

Despite the lack of static analysis, we attempt to fully specify the type system
of \Trilogy{} in hopes that someday it can be statically analyzed (an attempt at
(potentially optional) static (higher-kinded? dependent?) typing is on the long
term roadmap).

Worth mentioning is that the type system of \Trilogy{} spans all sub-languages
consistently. The syntax and behaviour of all data types are the same, no matter
where they are being used. The types are summarized below, with exact specifications
of the operators come later, in Syntax and Semantics (\S\ref{sec:syntax})

\subsection{Built-In Types}

There are 8 primitive data types in \Trilogy{}:

\begin{itemize}
    \item Never
    \item Unit
    \item Boolean
    \item Atom
    \item Character
    \item String
    \item Number
    \item Bits
\end{itemize}

\noindent
There are also 6 compound types, which may be constructed out of primitives
and other types:

\begin{itemize}
    \item Struct
    \item Tuple
    \item Array
    \item Set
    \item Record
    \item Iterator
\end{itemize}

\noindent
Defined items (and their locally defined counterparts) are also values which
have types:

\begin{itemize}
    \item Procedure and Procedure Closure
    \item Function and Function Closure
    \item Continuation
    \item Module and Module Function
\end{itemize}

\noindent
Notably there is currently no type for a rule at this time. First class rules
may come later, but their semantics are unclear as rules are currently considered
more of pure control flow and pattern matching, and not as values themselves.

Additionally, there are no explicit user defined types, only combinations of the
built in types. That makes this list a complete summary of all types to be seen
in a \Trilogy{} program.

By no means, however, is this the end of the story for \Trilogy{}'s type
system. Custom abstract data types, as well as further built-in primitive types
are, are on the list of things to implement once the language is in a functioning
state.

\subsubsection{Serialization}

Most values in \Trilogy{} support multiple serializations when converting
to values of String type which are supported by the language runtime itself.

The ``string'' representation is used when converting to a String for
the purposes of debug printing or constructing messages for users.
This is the representation used when converted to String via template
interpolation or direct type conversion. Not all types have a string
representation; particularly functions, lambdas, and similar cannot
be converted to String in this way.

The ``source'' representation is the same as the value's source code literal
representation and is used when serializing or deserializing data as code.
This is particularly useful when generating code or a data file to be used
later. Not all values have a source representation; transient values such
as iterators may be infinite and cannot be written as a source literal, so
do not have source representations. Library- or user-defined functions may
be converted to a source representation; running such printed code should
be done with care as it may have unexpected consequences.

Other common representations (e.g. JSON or YAML) are left
to be implemented as library code.

\subsubsection{Structural Types}
\label{sec:structuraltypes}

Structural types are types which have a ``physical'' structure, which may be
interpreted intuitively. Structural types may be destructured using pattern
matching, and are compared structurally (deeply) using the structural equality
operator (\op{==}).

\subsubsection{Reference Types}
\label{sec:referencetypes}

Reference types are those which are stored and passed by \emph{reference}. When
a value of a reference type is passed as an argument or a assigned to a variable, both
variables refer to the same \emph{instance} of that value. When dealing with
mutation of an instance with multiple references, modifications made to any
reference to that instances are reflected in all other references. This should be
intuitive to users of Javascript, which follows similar reference semantics.

When comparing reference types, there are two options. The referential equality
operator (\op{===)} will compare the values by reference first; that is, if
both values are of a reference type, they are considered equal if and only
if they are references to the same instance. If either value is not a reference
type, they will be compared structurally. In contrast, the structural equality
operator (\op{==}) will compare structurally first: if both values are structural
types, they will be compared structurally, otherwise they will be compared by
reference.

Some types are both structural and reference types.

\subsubsection{Notation}

Though \Trilogy{} does not currently support in-source type annotation, a
type annotation syntax is defined, intended to be used in documentation
(i.e.\ comments) and to aid in reasoning about semantics within this specification.

The individual names and representations of specific types are described
in their respective sections. In addition to those types, the type sum operator
(\op{|}) may be used to describe a value that may be of one of two types (chaining
to reach more than two types), as in \ty{A | B}.

Another ``type'' that occasionally comes up in dynamically typed languages is the
``any'' type. In \Trilogy{}, an ``any'' type is avoided in favour of type variables,
which may be left unconstrained to represent accepting any other type in its place.

\input{primitives.tex}

\input{compound-types.tex}

\input{semantic-types.tex}
