\subsection{Primitive Types}

There's not much interesting to be said about primitives. These types are
backed directly by counterparts in the interpreter's implementation language,
and so cannot be further broken down in exchange for performance closer to
that of native code on their basic operations.

All primitive types are structural types, no primitive types are reference
types.

\subsubsection{Never}

Never is a type with no values. While not particularly useful as a value,
Never is the type of a \kw{return} statement, as well as the type of
a \kw{yield} that is not \kw{resume}d, and of actions which terminate an
execution, such as the \kw{end} statement, among other things.

In type annotation or documentation, the Never type is written \ty{!}.
In the formal semantics throughout this document, Never is typically not
mentioned, but is similar to the usage of $\bot$ to represent the code
that never completes execution.

\subsubsection{Unit}

Unit is a type with a single-value, the literal \kw{unit}. Unit on its own
has no particular meaning, but it can be used to signal the lack of value.
In type annotation, the Unit type is written \ty{Unit}.

This is the type returned by functions or procedures which do not return
any other value. Within the standard library, Unit is used as the sentinel
value for the end of a list.

The string representation of Unit is \texttt{unit}.

\subsubsection{Boolean}

Boolean is a type with two values, the literals \kw{true} and \kw{false},
which act much like they do in every language I can think of.
In type annotation, the Boolean type is written \ty{Bool}.

Boolean values are the result of many comparisons (such as \op{lt} and
\op{gt}), can be manipulated by the Boolean operators (\op{\&\&}, \op{||},
and \op{!}) and are used by control flow constructs (\kw{if} and
\kw{while}).

The string representation of the Boolean values
\kw{true} and \kw{false} are \texttt{true} and \texttt{false},
respectively.

\subsubsection{Atom}

Atom is a type which typically exist in finite quantities. Any individual
Atom has no value on its own, but its literal form may hint to a reader at
what its value is meant to signal. The only real operation available for
Atom values is equality, which is performed in $O(1)$ time.

An Atom literal is an identifier prefixed with a single quote (\texttt{'}).
Within the standard library, some Atoms yielded as effects to signal
exceptional situations:

In type annotation, the Atom type is written \ty{Atom} when any Atom
is expected, or using the exact value of the atom (e.g. \ty{'NAN}) if
only that exact value is expected.

\begin{itemize}
    \item \val{'NAN}: yielded by Number parsing functions when the value
    cannot be parsed as a Number.
    \item \val{'EOF}: yielded by input procedures when the end of input
    has been reached.
\end{itemize}

The string representation of an Atom is its symbol without the single quote
(i.e. \val{'NAN} as a string is \texttt{NAN}).

\subsubsection{Character}

The Character type is used to represent individual Unicode code points,
a Character value representing exactly one.
In type annotation, the Character type is written \ty{Char}.

Internally, Characters are represented using their UTF-8 encodings. Characters
can safely be converted to Numbers, as every Character's UTF-8 encoding is a valid
Number. Meanwhile, not every Number may be safely converted to a Character,
as not every binary sequence represents a valid UTF-8 codepoint.

Character literals are specified as a single UTF-8 code point between
single quotes. Alternatively, an escape sequence (\S\ref{sec:escapesequences})
may be used to represent an ASCII or Unicode character code by its hexadecimal
representation, or one of the specially supported common escape sequences.

The string representation of a Character is a String containing that one
Character. Some examples:

\begin{table}[H]
    \centering
    \begin{tabular}{ll}
        \hline
        \textbf{Source} & \textbf{String} \\
        \hline
        \val{'c'} & \texttt{c} \\
        \val{'\textbackslash t'} & \emph{Line break}\\
        \val{'\textbackslash x65'} & \texttt{e} \\
        \val{'\textbackslash u\{03BB\}} & \texttt{$\lambda$} \\
        \hline
    \end{tabular}
\end{table}

\subsubsection{String}

The String type is a sequence of Characters. This sounds somewhat like
a compound type, but since a String can only contain Characters and has
its own literal format, it is practically more like a primitive. In type
annotation, the String type is written \ty{String}.

Strings may be manipulated using the push operator (\op{+}) to attach a
Character to a String, and the glue operator (\op{<>}) to attach two
Strings together. The member access operator (\op{.}) may also be used to
access a Character at a particular index.

String literals are represented using double quotes, and may contain
any characters or escape sequences. Untagged template literals produce
Strings by inserting the string representation of the interpolated values
(tagged literals don't actually produce Strings).

The string representation of a String is the string itself.

\subsubsection{Number}

The Number type is used to represent numbers, on which mathematical
operations can be performed. In \Trilogy{}, numbers go beyond integers
and floating point numbers to also include any rational or complex
numbers as well. In type annotation, the Number type is written \ty{Number}.
Additonally, the names \ty{Int}, \ty{Float}, and \ty{Rational} refer to
different subsets of the full range of the \ty{Number} type.

These numbers support all the normal operations (addition \op{+},
subtraction \op{-}, multiplication \op{*}, division \op{/},
modulus \op{\%}, integer division \op{//}, power \op{**}), but
division is perfect (\texttt{1 / 3 == 1/3}), and operations that
require complex numbers (\texttt{(-1) ** 1/2 == 0i1}) will succeed.

The string representation of a Number depends on the value of the number,
and is most conveniently described by example:

\begin{table}[H]
    \centering
    \begin{tabular}{rr}
        \hline
        \textbf{Source} & \textbf{String} \\
        \hline
        \val{0} & \texttt{0} \\
        \val{123} & \texttt{123} \\
        \val{-123} & \texttt{-123} \\
        \val{123.456} & \texttt{123.456} \\
        \val{-123.456} & \texttt{-123.456} \\
        \val{1/3} & \texttt{1/3} \\
        \val{-1/3} & \texttt{-1/3} \\
        \val{123i456} & \texttt{123 + 456i} \\
        \val{123 - 0i456} & \texttt{123 - 456i} \\
        \val{123.5i5.4} & \texttt{123.5 + 5.4i} \\
        \hline
    \end{tabular}
\end{table}

\noindent
There are some cases missing, but hopefully they are easy to extrapolate.

\subsubsection{Bits}

The Bits type is used to represent binary numbers as sequences of bits.
While many languages use their number type for this purpose, \Trilogy{}
differentiates Bits from Numbers due to the fact that Numbers are represented
in arbitrary precision and so bitwise operations (and \op{\&}, or \op{|},
xor \op{\textasciicircum}, not \op{\textasciitilde}, shift \op{\textasciitilde>}
and \op{<\textasciitilde}) do not work all that well on them.
In type annotation, the Bits type is written \ty{Bits}.

The string representation of Bits is the literal value of those bits in full
represented using 1s and 0s, as in \texttt{010101}. The source representation
is as the bits literal when written with prefix \texttt{0bb} as in as in
\texttt{0bb010101}
