\section{Lexical Structure}

Grammars in this document are typically
\href{https://en.wikipedia.org/wiki/Parsing_expression_grammar}{PEGs},
despite maybe looking like BNF (though, really, do they look any different?).
I did not find a particularly good grammar-typesetting package in a brief
Google search. Excuse any typographic inconsistencies within the grammars
at this time.

In this section, capitalized productions represent actual token types, the
names of which reflect the names of tokens found in the implementation source
code. Lowercase productions represent internal rules which are not seen in the
resulting token stream after lexing.

\subsection{Input Format}

\Trilogy{} input is interpreted as UTF-8 text. Any sequence of bytes which
does not constitute a valid UTF-8 codepoint sequence is not valid \Trilogy{}
code.

Including the Unicode \href{https://en.wikipedia.org/wiki/Byte_order_mark}{byte-order mark}
(BOM \texttt{U+FEFF}) is \emph{strongly recommended against}, but \Trilogy{}
implementations may accept its existence (and ignore it) with a warning.

A valid \Trilogy{} file should end with a new line character (\texttt{\textbackslash n}),
but \Trilogy{} implementations may accept a file that does not have
a trailing new line by implying the existence of one, and emitting a
warning.

\subsection{Keywords}
\FloatBarrier

Keywords in \Trilogy{} (listed in Table~\ref{tab:keywords}) have
different meanings in the different sub-languages, but \emph{all}
keywords are considered ``reserved'' words in all contexts. That
is, they may only be used as keywords, even in sub-languages which
do not have a use for that keyword. Usage as identifiers in any
situation are prohibited.

\begin{table}[h]
\parbox[t]{0.45\linewidth}{
    \centering
    \begin{tabular}[t]{ll}
        \hline
        \textbf{Token Type} & \textbf{Keyword} \\
        \hline
        \texttt{KwAnd} & \kw{and} \\
        \texttt{KwAssert} & \kw{assert} \\
        \texttt{KwAs} & \kw{as} \\
        \texttt{KwBecome} & \kw{become} \\
        \texttt{KwBreak} & \kw{break} \\
        \texttt{KwCancel} & \kw{cancel} \\
        \texttt{KwConst} & \kw{const} \\
        \texttt{KwCase} & \kw{case} \\
        \texttt{KwContinue} & \kw{continue} \\
        \texttt{KwDefer} & \kw{defer} \\
        \texttt{KwDo} & \kw{do} \\
        \texttt{KwElse} & \kw{else} \\
        \texttt{KwEnd} & \kw{end} \\
        \texttt{KwExit} & \kw{exit} \\
        \texttt{KwExport} & \kw{export} \\
        \texttt{KwExtern} & \kw{extern} \\
        \texttt{KwFalse} & \kw{false} \\
        \texttt{KwFn} & \kw{fn} \\
        \texttt{KwFor} & \kw{for} \\
        \texttt{KwFunc} & \kw{func} \\
        \texttt{KwIf} & \kw{if} \\
        \texttt{KwImport} & \kw{import} \\
        \texttt{KwIn} & \kw{in} \\
        \texttt{KwIs} & \kw{is} \\
        \hline
    \end{tabular}
}
\hfill
\parbox[t]{0.45\linewidth}{
    \centering
    \begin{tabular}[t]{ll}
        \hline
        \textbf{Token Type} & \textbf{Keyword} \\
        \hline
        \texttt{KwLet} & \kw{let} \\
        \texttt{KwMatch} & \kw{match} \\
        \texttt{KwMut} & \kw{mut} \\
        \texttt{KwNot} & \kw{not} \\
        \texttt{KwOr} & \kw{or} \\
        \texttt{KwPass} & \kw{pass} \\
        \texttt{KwProc} & \kw{proc} \\
        \texttt{KwQy} & \kw{qy} \\
        \texttt{KwResume} & \kw{resume} \\
        \texttt{KwReturn} & \kw{return} \\
        \texttt{KwRule} & \kw{rule} \\
        \texttt{KwSuper} & \kw{super} \\
        \texttt{KwTest} & \kw{test} \\
        \texttt{KwThen} & \kw{then} \\
        \texttt{KwTrue} & \kw{true} \\
        \texttt{KwType} & \kw{type} \\
        \texttt{KwTypeof} & \kw{typeof} \\
        \texttt{KwUnit} & \kw{unit} \\
        \texttt{KwUse} & \kw{use} \\
        \texttt{KwWhen} & \kw{when} \\
        \texttt{KwWhile} & \kw{while} \\
        \texttt{KwWith} & \kw{with} \\
        \texttt{KwYield} & \kw{yield} \\
        \hline
    \end{tabular}
}
\caption{\label{tab:keywords}\Trilogy{} Keywords}
\end{table}

\subsection{Reserved Words}

Some additional words (Table~\ref{tab:reservedwords}) are reserved for
future usage, following the same usage restrictions as keywords. This
list is intentionally very restrictive (who knows what ill advised feature
I might add to this language next) and will be relaxed over time as features
are ruled out or implemented as library code. Notably, this includes
most keywords I can think of from other languages.

\begin{table}[h]
\parbox[t]{0.45\linewidth}{
    \centering
    \begin{tabular}[t]{ll}
        \hline
        \textbf{Token Type} & \textbf{Keyword} \\
        \hline
        \texttt{KwAsync} & \kw{async} \\
        \texttt{KwAwait} & \kw{await} \\
        \texttt{KwCatch} & \kw{catch} \\
        \texttt{KwClass} & \kw{class} \\
        \texttt{KwData} & \kw{data} \\
        \texttt{KwEnum} & \kw{enum} \\
        \texttt{KwExcept} & \kw{except} \\
        \texttt{KwExtends} & \kw{extends} \\
        \texttt{KwImplements} & \kw{implements} \\
        \texttt{KwInline} & \kw{inline} \\
        \texttt{KwInstanceof} & \kw{instanceof} \\
        \texttt{KwInterface} & \kw{interface} \\
        \texttt{KwIter} & \kw{iter} \\
        \texttt{KwLazy} & \kw{lazy} \\
        \texttt{KwLens} & \kw{lens} \\
        \texttt{KwLoop} & \kw{loop} \\
        \hline
    \end{tabular}
}
\hfill
\parbox[t]{0.45\linewidth}{
    \centering
    \begin{tabular}[t]{ll}
        \hline
        \textbf{Token Type} & \textbf{Keyword} \\
        \hline
        \texttt{KwMacro} & \kw{macro} \\
        \texttt{KwModule} & \kw{module} \\
        \texttt{KwNext} & \kw{next} \\
        \texttt{KwOper} & \kw{oper} \\
        \texttt{KwPrec} & \kw{prec} \\
        \texttt{KwProtocol} & \kw{protocol} \\
        \texttt{KwStatic} & \kw{static} \\
        \texttt{KwStruct} & \kw{struct} \\
        \texttt{KwSwitch} & \kw{switch} \\
        \texttt{KwTag} & \kw{tag} \\
        \texttt{KwTrait} & \kw{trait} \\
        \texttt{KwTry} & \kw{try} \\
        \texttt{KwUnless} & \kw{unless} \\
        \texttt{KwUntil} & \kw{until} \\
        \texttt{KwWhere} & \kw{where} \\
        \hline
    \end{tabular}
}
\caption{\label{tab:reservedwords}\Trilogy{} Reserved Words}
\end{table}

\FloatBarrier
\subsection{Identifiers}

Identifiers in \Trilogy{} are currently restricted to US-ASCII characters,
with the intent being to support Unicode identifiers (including Emoji) in
a later edition. As a special case, an identifier ending with an equal
sign (\texttt{=}) is parsed as a single token for functions when used
in assignment operator position.

An additional restriction: the single underscore (\texttt{\_}) is never an
identifier, and is instead classified separately.

\begin{bnf*}
    \bnfprod{Discard}{\bnfts{\_}}\\
    \bnfprod{IdentifierEq}{\bnfpn{id}\bnfsp\bnfts{=}}\\
    \bnfprod{Identifier}{\bnfpn{id}\quad\bnftd{excluding keywords}}\\
    \bnfprod{id}{\bnfts{[a-zA-Z\_][a-zA-Z0-9\_]*}} \\
\end{bnf*}

\subsection{Atom Literals}

Atoms are represented as identifiers with a leading single-quote (\texttt{'}).

\begin{bnf*}
    \bnfprod{Atom}{\bnfts{'}\bnfsp\bnfpn{Identifier}} \\
\end{bnf*}

\subsection{Numeric Literals}

Numbers in \Trilogy{} are real numbers, as in supporting arbitrary
precision and largeness. Complex numbers are supported as well
(not that I have ever required the use of complex numbers in any
piece of code) for completeness.

Numeric tokens are of the following structure:

\begin{bnf*}
    \bnfprod{Numeric}{\bnfpn{real}\bnfor\bnfpn{real}\bnfsp\bnfpn{complex}} \\
    \bnfprod{real}{\bnfpn{integer}\bnfor\bnfpn{rational}\bnfor\bnfpn{float}} \\
    \bnfprod{integer}{
        \bnfpn{decimal}
        \bnfor
        \bnfpn{hexadecimal}
        \bnfor
        \bnfpn{binary}
        \bnfor
        \bnfpn{octal}
    } \\
    \bnfprod{decimal}{\bnfts{[0-9][0-9\_]*}} \\
    \bnfprod{hexadecimal}{\bnfts{0x}\bnfsp\bnfts{[0-9A-Fa-f][0-9A-Fa-f\_]*}} \\
    \bnfprod{binary}{\bnfts{0b}\bnfsp\bnfts{[01][01\_]*}} \\
    \bnfprod{octal}{\bnfts{0o}\bnfsp\bnfts{[0-7][0-7\_]*}} \\
    \bnfprod{rational}{\bnfpn{integer}\bnfsp\bnfts{/}\bnfsp\bnfpn{integer}} \\
    \bnfprod{float}{\bnfpn{decimal}\bnfsp\bnfts{.}\bnfsp\bnfpn{decimal}} \\
    \bnfprod{complex}{\bnfts{i}\bnfsp\bnfpn{real}}
\end{bnf*}

Negative numbers do not exist in source code, and are actually
an appliction of the unary negation operator \op{-} to a non-negative number.
Numbers with exponential components also do not exist.

Numbers may contain the underscore character (\texttt{\_}) as a separator.
There are no restrictions on the placement of underscores, beyond that they
may not start a number. \val{1000000}, \val{1\_000\_000}, and
\val{1\_0\_0\_\_\_000\_0} are all valid number literals representing one
million.

Worth noting as well, rational number literals (e.g. \texttt{2/3}) are
parsed as a single token, while runtime division (e.g. \texttt{2 / 3})
is parsed as three separate tokens. Be careful of this distinction later
when writing or parsing expressions, as these two representations will
lead to different results in ways not relevant to other languages without
rational literals.

As they are not often seen in other languages (and similarly unlikely to
be witnessed in \Trilogy{}), complex numbers are written such as \val{13i5}
to represent $13 + 5i$. On either side of the \texttt{i} may be a valid
real number (including rational or floating point numbers). Neither value
may be omitted (i.e. \val{0i5} not \texttt{i5}, and \val{13i0} not
\texttt{13i}). A negative complex number $15 - 5i$ is written \val{15 - 0i5}
(there is no literal, this is just a subtraction).

\subsection{String and Character Literals}

Strings in \Trilogy{} are always enclosed in double quotes (\texttt{"}).

Special ``template strings'' are prefixed with a \texttt{\$}, or tagged
template strings are prefixed with a \texttt{\$} and an identifier.
String tokenization ends up producing many types of tokens for template
strings, where template strings containing no interpolations are parsed
as regular strings:

\begin{bnf*}
    \bnfprod{String}{
        \bnfts{"}
        \bnfsp
        \bnfpn{content}
        \bnfsp
        \bnfts{"}
    } \\
    \bnfprod{DollarString}{
        \bnfts{\$"}
        \bnfsp
        \bnfpn{content}
        \bnfsp
        \bnfts{"}
    }\\
    \bnfprod{TaggedString}{
        \bnfts{\$"}
        \bnfsp
        \bnfpn{content}
        \bnfsp
        \bnfts{"}
    }\\
    \bnfprod{TemplateStart}{
        \bnfts{\$"}
        \bnfsp
        \bnfpn{content}
        \bnfsp
        \bnfts{\$\{}
        \bnfor
    } \\
    \bnfmore{
        \bnfts{\$}
        \bnfsp
        \bnfts{Identifier}
        \bnfsp
        \bnfts{"}
        \bnfsp
        \bnfpn{content}
        \bnfsp
        \bnfts{\$\{}
        \bnfor
    } \\
    \bnfprod{TemplateMiddle}{
        \bnfts{\}}
        \bnfsp
        \bnfpn{content}
        \bnfsp
        \bnfts{\$\{}
    } \\
    \bnfprod{TemplateEnd}{
        \bnfts{\}}
        \bnfsp
        \bnfpn{content}
        \bnfsp
        \bnfts{"}
    } \\
    \bnfprod{content}{\bnftd{Any UTF-8 characters or escape sequences}} \\
\end{bnf*}

Similarly, characters are always enclosed in single quotes (\texttt{'}).
Unlike many other dynamically typed languages, \Trilogy{} \emph{does} differentiate
between characters and strings. Character literals may contain either a
single (non-single-quote) character, or a single escape sequence.

\begin{bnf*}
    \bnfprod{Character}{\bnfts{'}\bnfsp\bnfpn{char}\bnfsp\bnfts{'}} \\
    \bnfprod{char}{\bnftd{One Unicode character}\bnfor\bnftd{One escape sequence}} \\
\end{bnf*}

\subsubsection{Escape Sequences}
\label{sec:escapesequences}
\FloatBarrier

Within character and string literals, escape sequences may be used to denote
special characters, typically ones that would otherwise be hard to physically
type or print and read. Valid escape sequences and their meanings are listed
in Table~\ref{tab:escape}.

\begin{table}[h]
    \centering
\begin{tabular}[c]{ll}
    \hline
    \textbf{Code} & \textbf{Usage} \\
    \hline
    \texttt{\textbackslash n} & New line \\
    \texttt{\textbackslash r} & Carriage Return \\
    \texttt{\textbackslash t} & Tab \\
    \texttt{\textbackslash\textbackslash} & Backslash \\
    \texttt{\textbackslash 0} & Null \\
    \texttt{\textbackslash '} & Single Quote \\
    \texttt{\textbackslash "} & Double Quote \\
    \texttt{\textbackslash \$} & Dollar Sign \\ %
    \texttt{\textbackslash x41} & ASCII escape \\
    \texttt{\textbackslash u\{7FFF\}} & Unicode escape sequence \\
    \hline
\end{tabular}
\caption{\label{tab:escape}Escape Sequences}
\end{table}

Escape sequences in both strings and characters are resolved by the lexer
and replaced before the tokens are handed off to further passes.

\FloatBarrier
\subsection{Bits Literals}

\Trilogy{} supports a specific type for arbitrary binary strings, distinct
from numbers. These values are written similarly to integers, but with an
extra ``b''. As with numeric literals, bits literals may use \texttt{\_}
as a separator.

\begin{bnf*}
    \bnfprod{Bits}{
        \bnfpn{bitshexadecimal}
        \bnfor
        \bnfpn{bitsbinary}
        \bnfor
        \bnfpn{bitsoctal}
    } \\
    \bnfprod{bitshexadecimal}{\bnfts{0bx}\bnfsp\bnfts{[0-9A-Fa-f][0-9A-Fa-f\_]*}} \\
    \bnfprod{bitsbinary}{\bnfts{0bb}\bnfsp\bnfts{[01][01\_]*}} \\
    \bnfprod{bitsoctal}{\bnfts{0bo}\bnfsp\bnfts{[0-7][0-7\_]*}} \\
\end{bnf*}

\subsection{Whitespace}

While \Trilogy{} is a mostly free-form language; indentation levels do not affect
the meaning of the code. Whitespace does remain signficant in a few places, slightly
more intrusively than in some other languages. In particular, certain parts of the
code may be parsed differently depending on whether the tokens were flush, spaced, or
on separate lines. For this reason, spaces, tabs, and line feeds are emitted by the
scanner as distinct tokens.

\begin{bnf*}
    \bnfprod{Space}{
        \bnfts{U+0009}
        \bnfsp
        \bnfpn{Space}
        \bnfor
        \bnfts{U+0020}
        \bnfsp
        \bnfpn{Space}
        \bnfor
    } \\
    \bnfmore{
        \bnfts{U+0009}
        \bnfor
        \bnfts{U+0020}
    } \\
    \bnfprod{EndOfLine}{\bnfts{U+000A}}
\end{bnf*}

The carriage return character is permitted in the source file, so as to support both
LF and CRLF style line endings, but is not treated as whitespace. A carriage return
anywhere except directly beside it's corresponding line feed is unlikely.

Specifically, the four whitespace characters permitted in \Trilogy{} source code
are the following:

\begin{itemize}
    \item \texttt{U+0009} (tab \texttt{'\textbackslash t'})
    \item \texttt{U+000A} (line feed \texttt{'\textbackslash n'})
    \item \texttt{U+000D} (carriage return \texttt{'\textbackslash r'})
    \item \texttt{U+0020} (space \texttt{' '})
\end{itemize}

\subsection{Comments}

Comments hold no semantic meaning, but they may constitute a line break,
if there is a line break within the comment. A line comment continues until and
includes the line feed character, and so are always considered as line breaks.
Comments which do not contain a line feed are treated as a non-line
break whitespace. Block comments may be nested.

To distinguish between line-breaking comments and non-line breaking comments,
block comments have two different token types.

\begin{bnf*}
    \bnfprod{CommentInline}{
        \bnfts{\#-}
        \bnfsp
        \bnftd{No line feed}
        \bnfsp
        \bnfpn{CommentInline}
        \bnfsp
        \bnftd{No line feed}
        \bnfsp
        \bnfts{-\#}
        \bnfor
    } \\
    \bnfmore{\bnfts{\#-} \bnfsp \bnftd{No line feed} \bnfsp \bnfts{-\#}} \\
    \bnfprod{CommentBlock}{
        \bnfts{\#-}
        \bnfsp
        \bnftd{Anything}
        \bnfsp
        \bnfpn{CommentBlock}
        \bnfsp
        \bnftd{Anything}
        \bnfsp
        \bnfts{-\#}
        \bnfor
    } \\
    \bnfmore{\bnfts{\#-} \bnfsp \bnftd{Anything} \bnfsp \bnfts{-\#}} \\
    \bnfprod{DocOuter}{\bnfts{\#\#} \bnfsp \bnftd{All characters until line feed}} \\
    \bnfprod{DocInner}{\bnfts{\#!} \bnfsp \bnftd{All characters until line feed}} \\
    \bnfprod{CommentLine}{\bnfts{\#} \bnfsp \bnftd{All characters until line feed}} \\
\end{bnf*}

\subsection{Punctuation}
\FloatBarrier

Punctuation, listed in Table~\ref{tab:operators}, includes symbols and operators that
treated as whole tokens when encountered on their own.

Keep in mind the slash (\texttt{/}) also appears as part of rational number literals, so
will not be exclusively identified as an individual token.

Not all tokenized punctuation marks are used; a few are reserved for future, or tokenized
only to produce clearer error messages.

\begin{table}[h]
\parbox[t]{0.45\linewidth}{
    \centering
    \begin{tabular}[t]{ll}
        \hline
        \textbf{Token Type} & \textbf{Symbol} \\
        \hline
        \texttt{OpEq} & \op{=} \\
        \texttt{OpLt} & \op{<} \\
        \texttt{OpGt} & \op{>} \\
        \texttt{OpEqEq} & \op{==} \\
        \texttt{OpBangEq} & \op{!=} \\
        \texttt{OpEqEqEq} & \op{===} \\
        \texttt{OpBangEqEq} & \op{!==} \\
        \texttt{OpLtEq} & \op{<=} \\
        \texttt{OpGtEq} & \op{>=} \\
        \hline
        \texttt{OpBang} & \op{!} \\
        \texttt{OpAmpAmp} & \op{\&\&} \\
        \texttt{OpPipePipe} & \op{||} \\
        \texttt{OpAmpAmpEq} & \op{\&\&=} \\
        \texttt{OpPipePipeEq} & \op{||=} \\
        \hline
        \texttt{OpAmp} & \op{\&} \\
        \texttt{OpPipe} & \op{|} \\
        \texttt{OpCaret} & \op{\textasciicircum} \\
        \texttt{OpTilde} & \op{\textasciitilde} \\
        \texttt{OpShr} & \op{\textasciitilde>} \\
        \texttt{OpShl} & \op{<\textasciitilde} \\
        \texttt{OpShrEx} & \op{\textasciitilde\textasciitilde>} \\
        \texttt{OpShlEx} & \op{<\textasciitilde\textasciitilde} \\
        \texttt{OpShrCon} & \op{\textasciitilde>>} \\
        \texttt{OpShlCon} & \op{<<\textasciitilde} \\
        \hline
        \texttt{OpAmpEq} & \op{\&=} \\
        \texttt{OpPipeEq} & \op{|=} \\
        \texttt{OpCaretEq} & \op{\textasciicircum=} \\
        \texttt{OpShrEq} & \op{\textasciitilde>=} \\
        \texttt{OpShlEq} & \op{<\textasciitilde=} \\
        \texttt{OpShrExEq} & \op{\textasciitilde\textasciitilde>=} \\
        \texttt{OpShlExEq} & \op{<\textasciitilde\textasciitilde=} \\
        \texttt{OpShrConEq} & \op{\textasciitilde>>=} \\
        \texttt{OpShlConEq} & \op{<<\textasciitilde=} \\
        \hline
        \texttt{OpColonColon} & \op{::} \\
        \hline
    \end{tabular}
}
\hfill
\parbox[t]{0.45\linewidth}{
    \centering
    \begin{tabular}[t]{ll}
        \hline
        \textbf{Token Type} & \textbf{Symbol} \\
        \hline
        \texttt{OpGlue} & \op{<>} \\
        \texttt{OpGlueEq} & \op{<>=} \\
        \texttt{OpColon} & \op{:} \\
        \texttt{OpColonEq} & \op{:=} \\
        \hline
        \texttt{OpPlus} & \op{+} \\
        \texttt{OpMinus} & \op{-} \\
        \texttt{OpStar} & \op{*} \\
        \texttt{OpSlash} & \op{/} \\
        \texttt{OpSlashSlash} & \op{//} \\
        \texttt{OpPercent} & \op{\%} \\
        \texttt{OpStarStar} & \op{**} \\
        \hline
        \texttt{OpPlusEq} & \op{+=} \\
        \texttt{OpMinusEq} & \op{-=} \\
        \texttt{OpStarEq} & \op{*=} \\
        \texttt{OpSlashEq} & \op{/=} \\
        \texttt{OpSlashSlashEq} & \op{//=} \\
        \texttt{OpPercentEq} & \op{\%=} \\
        \texttt{OpStarStarEq} & \op{**=} \\
        \hline
        \texttt{OpDot} & \op{.} \\
        \texttt{OpDotEq} & \op{.=} \\
        \texttt{OpDotDot} & \op{..} \\
        \texttt{OpComma} & \op{,} \\
        \texttt{OpSemi} & \op{;} \\
        \texttt{OpLeftArrow} & \op{<-} \\
        \texttt{OpFatArrow} & \op{=>} \\
        \hline
        \texttt{OpLtLt} & \op{<<} \\
        \texttt{OpGtGt} & \op{>>} \\
        \texttt{OpLtLtEq} & \op{<<=} \\
        \texttt{OpGtGtEq} & \op{>>=} \\
        \texttt{OpPipeGt} & \op{|>} \\
        \texttt{OpLtPipe} & \op{<|} \\
        \hline
        \texttt{OpQuestion} & \op{?} \\
        \texttt{OpTildeEq} & \op{\textasciitilde=} \\
        \texttt{OpRightArrow} & \op{->} \\
        \hline
    \end{tabular}
}
\caption{\label{tab:operators}Operators and Punctuation}
\end{table}

\subsection{Delimeters}

Delimiters, listed in Table~\ref{tab:delimiters} typically surround other
pieces of code, and must come in matched opening-closing pairs.

\begin{table}[h]
    \centering
    \begin{tabular}[t]{llll}
        \hline
        \textbf{Token Type} & \textbf{Symbol}& \textbf{Token Type} & \textbf{Symbol} \\
        \hline
        \texttt{OBrace} & \texttt{\{}& \texttt{CBrace} & \texttt{\}} \\
        \texttt{OBracePipe} & \texttt{\{|}& \texttt{CBracePipe} & \texttt{|\}} \\
        \texttt{OBrack} & \texttt{[} & \texttt{CBrack} & \texttt{]} \\
        \texttt{OBrackPipe} & \texttt{[|}& \texttt{CBrackPipe} & \texttt{|]} \\
        \texttt{OParen} & \texttt{(} & \texttt{CParen} & \texttt{)} \\
        \hline
    \end{tabular}
    \caption{\label{tab:delimiters}Delimiters}
\end{table}

\subsection{Shebang}

A \Trilogy{} file may optionally begin with a
\href{https://en.wikipedia.org/wiki/Shebang_(Unix)}{shebang}. This has no bearing
on how the program is run or interpreted when provided to a \Trilogy{} implementation,
but may affect how the operating system treats the file.

Conveniently, this matches standard comment syntax. Inconveniently, it is not
likely that it is meant to be interpreted as a doc comment. Conveniently,
we don't care, and so it is up to the documentation generator to detect and
hide a shebang from generated documentation if so desired.

\FloatBarrier
