\subsection{Poetry}

As \Law{} is a language of binding, and \Prose{} a language of control
flow, \Poetry{} is a language of value. As a pure functional programming
language, \Poetry{} is highly expression-based, though given the existence
of the effect system it is possible to trigger side effects, in a way.

\begin{bnf*}
    \bnfprod{FunctionHead}{
        \bnfpn{Identifier}
        \bnfsp
        \bnfpn{ParameterList}
    } \\
    \bnfprod{ParameterList}{
        \bnfpn{Pattern}
        \bnfsp
        \bnfpn{PatternList}
        \bnfor
        \bnfpn{Pattern}
    } \\
    \bnfprod{FunctionBody}{\bnfpn{Expression}}
\end{bnf*}

Similar to Haskell, functions in \Poetry{} are defined as equations. Defining
the same name multiple times is allowed, given that its number of arguments
is the same, and the patterns of its parameters are changed, allowing different
cases of a function to be handled by different expressions. When the function is
applied, the first matching definition in source code order is used, and the
others are skipped.

Also similar to Haskell, functions support currying and partial application by
default. In \Trilogy{}, this is defined as a syntax transformation over function
definitions with multiple parameters into unary functions returning a function
that accepts the next parameter.

Unsimilarly to Haskell, \Poetry{} is evaluated eagerly. This means that nullary
(0-argument) functions are pretty useless, so all functions \emph{must} have a
parameter. Nullary procedures may be a simple alternative in the meantime, where
the calling of that procedure is explicit, or a function accepting a single
ignored \kw{unit} parameter, if truly \Poetry{} is preferred.

\begin{align*}
    \texttt{\$f:id \$(\$r:pat)+ \$p:pat = \$e:expr} & \Rightarrow \texttt{\$f \$(\$r)+ = \kw{fn} \$p. \$e} %
\end{align*}

%% TODO: I really can't decide if it's worth making Poetry lazy. Technically it is
%% a pretty neat thing to be able to work in lazy code, but practically... I find
%% it less so. Poetry is meant to be practical.
%%
%% This does eliminate the ability to have nullary functions, but I really don't
%% mind. You can just pass a unit anyway.

% The parameter list may be empty, defining a nullary (0-argument) function. As
% \Poetry{} is lazy, this function will only be evaluated when its result is
% needed, so even a nullary function may, for example, yield an effect which will
% only be performed when it is required. Note that the lazy evaluation of \Poetry{}
% can lead to some code being run in unexpected order, so it is recommended to
% think carefully about whether effects in \Poetry{} are required.

% Though \Poetry{} is lazy, \Prose{} and \Law{} (except for through Iterators)
% are not. When an expression of \Poetry{} appears in \Prose{} or \Law{},
% as in an assignment or unification, it is evaluated immediately. This implies
% that while \Poetry{} may be used to represent infinite data structures those
% data structures must remain in \Poetry{} or else they will be infinitely
% evaluated, unless converted to Iterator.

\begin{prooftree}
    \def\defaultHypSeparation{\hskip 0.1in}
    \AxiomC{$\kw{func}\ f\ P\ \texttt{=}\ E$}
    \AxiomC{$\Sigma\vdash X : \tau$}
    \AxiomC{$P = X$}
    \AxiomC{$\Gamma,\setof{a\given a\defby P}\vdash E : \rho $}
    \LeftLabel{Function}
    \QuaternaryInfC{$\Gamma\vdash f : \tau \rightarrow \rho$}
\end{prooftree}

\input{expressions.tex}

\input{literal.tex}

\input{comprehension.tex}

\input{reference.tex}

\input{member-access.tex}

\input{application.tex}

\input{procedure-call.tex}

\input{binary-op.tex}

\input{unary-op.tex}

\input{let.tex}

\input{match.tex}

\input{is.tex}

\input{end.tex}

\input{first-class-keyword.tex}

\input{closures.tex}

\input{template.tex}

\input{precedence.tex}
